Best Practices for Alchemical Free Energy Calculations: v\versionnumber

### 2019-12-10 paper/manuscript.tex
most existing references now fixed

Below, we provide a non-exhaustive list of topics we have not addressed with some references to provide starting points on these more advanced topics:
\begin{itemize}
\item Covalent inhibition.
\item Complex association stoichiometry (beyond simple 1:1 association)~\cite{awesome reference}. 
\item Endpoint free energy methods such as MM-PBSA~\cite{genheden2015mm} and LIE~\cite{utierrez-de-teran2012lineara}
\item Free energy methods that extract the ligand using geometric order parameters and potential of mean force methods~\cite{heinzelmann2017attachpullrelease}
\item Forcefield dependence for protein, ligand, ions, co-solvents, and co-factors. Different studies have looked at the influence of force fields and it is assumed the user has made an adequate choice for the system under study.~\cite{loeffler2018reproducibility, vassetti2019assessment, lopes2015current} 
\end{itemize}

@@

How can we compute $\Delta G_{\mathrm{bind}}$ from a molecular dynamics simulation? Running a direct simulation of the binding events is too computationally costly, meaning we cannot easily estimate the configurational integrals of the bound and unbound state from the simulation. This is where alchemical free simulations come in. We can write down a thermodynamic cycle as seen in ~\ref{fig:illustrative-alchemical-transformation}, for a relative free energy simulation. Based on this cycle we can compute the relative free energy of binding between two compounds $A$ and $B$, by going along the arrow of $\Delta G_{\mathrm{bound}}$, where we introduce a second compound that is similar to the first one. Now $\Delta G_{\mathrm{bound}}$ can be expressed as a ratio of partition functions of toluene (A) and benzyl alcohol (B). 

@@

Current alchemical free energy calculations seem to achieve, in favorable cases, RMS errors around 1-2 kcal/mol depending on force field, system, and a variety of other factors such as simulation time, sampling method, and whether the calculations employed are absolute or relative. An overview of benchmark datasets and case studies can be found in section \ref{sec:benchmark} at the end of this document.
\todo[inline,color={green!20}]{JDC: We should probably provide a short list of key benchmark references here, broken down by force field, and keep this up to date. GT: would a table, at the end of the document be an easier way to do this, only requiring some brief comments in the text? ASJSM: I like the table idea. MK: Feel free to remove this todo item if you are happy with the table provided. I refrained from just listing lots of case studies, as I think we should highlight the recommended datasets which should be used for benchmarking. Nevertheless, I have cited an review article containing such a summary as well as a recent publication comparing FEP+ and Gromacs in the brief text below the table.}
However, the domain of applicability is a significant concern~\cite{Sherborne2016, Cournia2017}, especially for relative calculations, which typically require a bound structure of a closely
related ligand as a starting point. Additional factors such as slow protein or ligand rearrangements, uncertainties in ligand binding mode, or charged ligands can make these calculations far less reliable and more of a research effort.

@@

If the binding site location is uncertain but the structure of the receptor is well defined and plausible binding sites are identified, you are more likely to chose an absolute free energy protocol to compute the standard free energy of binding of the ligand to a set of binding sites. This requires the user to prepare input files describing the bound conformation in different putative binding sites (REF). The apparent binding free energy of the ligand may be obtained by combining the individual binding site free energies, which also indicate where the ligand is more likely to bind. In this case using a docking program to generate initial structures is the way to go. Different commercial and none commercial tools are available, such as: rDock, Autodock Vina, Glide, or Flare to name a few~\cite{ruiz-carmona2014rdock, trott2010autodock, friesner2004glide, Flare}. 
If the putative binding sites are not apparent because of for instance significant induced-fit effects it may be challenging to obtain meaningful free energies of binding. One would have to account for the free energy change for forming a binding site in the target receptor. However, it may still be possible to obtain useful binding free energy estimates that may be compared between different ligands.

@@

\paragraph{Topologies} A critical first step in relative calculations is to select an approach to these calculations, determining whether to use a \emph{dual topology}, \emph{single topology}, or \emph{hybrid topology} approach to relative calculations.
\begin{figure}
    \includegraphics[width=0.95\columnwidth]{paper/figures/fig4_topol/Fig4.pdf}
    \caption{top: single topology bottom: dual topology Figure adapted from \url{http://www.alchemistry.org/wiki/Constructing_a_Pathway_of_Intermediate_States}}
    \label{fig:fig4_topology}
\end{figure} 
The distinction between these can be illustrated by considering a hypothetical transformation from molecule A to molecule B, where both atoms share a common substructure but differ in which functional groups are present; e.g. consider a transformation of benzene (A) to benzyl alcohol (B)~\ref{fig:topology}.
In this case the common substructure is the benzene ring, though perhaps may be larger depending on how it is defined, as we discuss below.
In single topology calculations, the overall transformation is set up to involve as few additional atoms as possible, so benzene would be typically changed into benzyl alcohol by changing one of the hydrogens into a carbon. To this new carbon you also find non-interacting atoms called ``dummy atoms'' (retaining their bonded interactions but not interacting with the rest of the system), which will become the two hydrogens attached to the new carbon and the OH group. Bond parameters as well as partial charges are adjusted accordingly. 
Thus in a single topology calculation, atoms may change their type so relatively few dummy atoms are created. This is illustrated in fig.~\ref{fig:fig4_topology} (a). 
In contrast, in a dual topology free energy calculation, no atoms are allowed to change type~\cite{shirts2012best}. This means that the benzene to benzyl alcohol transformation involves starting with benzene plus the non-interacting dummy atoms making up the hydroxy methyl group, then passing through an intermediate state where atoms which are becoming dummy atoms or ceasing being dummy atoms are partially interacting (this state may or may not be well defined~\cite{mobley2014blind}  and culminating in a state where benzyl alcohol is present along with the additional dummy atoms which was previously a corresponding hydrogen of the benzene. Fig.~\ref{fig:topology} (b) depicts how such a dual topology works.

@@

There are different tools that allow the generation of MCSS matches as well as single topology input. A large number of software tools are nowadays available for computing MCSS matches, some of the rely on RDKit for this computation, such as LOMAP or FESetup and partially BioSimSpace, others such as fkckombu are standalone tools~\cite{rdkit,loeffler2015fesetup, liu2013lead, hedges2019biosimspace, kawabata20143d}. Schr\"{o}dinger's FEP+ planning tool is based on a version of LOMAP, also allows MCSS matching and planning single topology calculations between molecules~\cite{wang2015accurate}.

@@

In binding free energy calculations, extra care should be taken when simulating the bound state, especially when dealing with weak binders and absolute free energy calculations in combination with enhanced sampling techniques.
In principle, only configurations of the receptor-ligand complex that we consider "bound" should be sampled from the bound state, and, in atomistic simulations, this requires to establish a structural definition of the bound state.
In practice, the simulation of the bound state starts with the ligand already placed in the binding site and relies on kinetic trapping to maintain a bound complex.
However, this strategy may not be sufficient if the complex dissociation rate is high or if methodologies such as Hamiltonian replica exchange~\ref{bitetti-putzer2003generalized} and expanded ensemble~\cite{chodera2011replica} are employed in absolute free energy calculations to enhance sampling since, in both cases, the ligand may find a way out of the binding site on timescales that are achievable by modern MD simulations.
A solution commonly adopted in these cases is the use of one or more restraints making the unbound configurations energetically unfavorable through the addition of extra terms in the potential function.
In practice, even with restraints, it is not always trivial to force a molecular dynamics simulation to explore a restricted region of the configurational space with a complicated geometry such as a binding site.
If the ligand is a tight binder\todo[inline, color={red!40}]{AR: Should we give an order of magnitude to define a tight and a weak binder?}, it is usually safe to employ a restraint that allows some of the unbound configurations to be sampled since they generally contribute negligibly to the partition function (i.e., they have a relatively small Boltzmann weight) as long as the sampled volume is not so large that their cumulative contribution becomes significant.
However, this is more problematic when dealing with weak binders as the unbound configurations can have a non-negligible Boltzmann weight, and their binding affinity can exhibit a significant dependency on the restraint type and parameters that are used to determine the sampled volume~\cite{}.
Moreover, it should be noted that the additional potential energy terms used to model the restraints can introduce a bias in the predicted free energy.
The bias can be removed at the analysis stage through reweighting techniques, but this procedure can increase the statistical uncertainty of the binding free energy estimate when the restraints are so strong that the overlap with the reweighted state is diminished~\cite{}.

@@

In a discovery setting, new ligands typically have unknown or at least uncertain binding modes~\cite{Kaus:2015:J.Chem.TheoryComput., PlountPrice:2000:J.Am.Chem.Soc.} [refs, including Mobley Structure paper and recent paper on non-additivity], complicating binding free energy estimation.
This uncertainty is because it is usually not desirable to estimate a binding affinity for a ligand which already has an available bound structure, since such a compound has already been tested.
To deal with prospective ligands with unknown binding modes, discovery projects commonly assume that modifications of functional groups on a common scaffold result in a consistent binding mode across all members of a series.
This is not necessarily always the case~\cite{Kaus:2015:J.Chem.TheoryComput.}, as reviewed elsewhere [ref Mobley Structure paper] and in some cases unexpected binding mode changes can be the origin of apparent non-additivity in structure-activity relationships [ref non-additivity paper].
Binding modes also tend to be particularly variable in the case of fragments, which often may have multiple relevant binding modes~\cite{}.

@@

In some cases, it is tempting to simply use docking techniques to generate initial bound structures for starting molecular dynamics simulations.
However, timescales for binding mode interconversion are usually slow compared to MD/free energy timescales, meaning that simulations started from different potential binding modes are likely to yield disparate computed binding free energies~\cite{Mobley:2006:TheJournalofChemicalPhysics, Palma:2012:J.Comput.Chem., Mobley:2012:TheJournalofChemicalPhysics, Gill:2018:J.Phys.Chem.B} .
And docking techniques are good at identifying sterically reasonable potential binding modes, but still perform relatively poorly at identifying a single dominant binding mode \emph{a priori}.~\cite{}

It is worth highlighting a recent SAMPL blind challenge on HIV integrase as an illustration of this. 
Many submissions, using state-of-the-art methods, had difficulty even predicting which \emph{binding site} ligands would bind in (most submissions placed more than half of the ligands into the incorrect binding site), and even given correct binding sites, the binding mode within each site was also quite difficult to predict~\cite{Mobley:2014:J.Comput.AidedMol.Des.}.
The best performing submission for predicting binding modes actually ended up being a human expert (aided by computational tools) with more than 10 years of experience on the particular target~\cite{Voet:2014:JournalofComputer-AidedMolecularDesign}, rather than a fully automated approach.
While free energy calculations on this set had some success, many of the failures actually ended up being cases where the binding mode selected as input for free energy calculations was later found to be incorrect~\cite{Gallicchio:2014:JournalofComputer-AidedMolecularDesign}, highlighting the importance of these issues.

One approach which has shown some success is to retain diverse potential binding modes from docking, perform short MD simulations of these to identify distinct stable binding modes, and then consider these in subsequent calculations~\cite{Gallicchio:2014:JournalofComputer-AidedMolecularDesign, Mobley:2006:TheJournalofChemicalPhysics, Rocklin:2013:JournalofMolecularBiology, Boyce:2009:JournalofMolecularBiology, Mobley:2007:JournalofMolecularBiology}.

Routes to handle multiple potential binding modes are different depending on whether absolute or relative calculations are selected, unless a method is available to estimate the relative populations of different stable binding modes in advance (e.g. such as the BLUES approach currently in development~\cite{Gill:2018:J.Phys.Chem.B}), in which case this approach could be applied to assist both types of calculations.

\paragraph{Handling multiple potential binding modes within absolute calculations.}
Within absolute binding free energy calculations, multiple potential binding modes can be handled by two main strategies: Consider each binding mode separately (a separation of states strategy) or sample all binding modes within a single simulation~\cite{Mobley:2012:TheJournalofChemicalPhysics}.
This couples to the choice of restraints selected, as some restraints will allow transitions between binding modes and even binding sites (Section~\ref{sec:standardstate-restraints}), and others do not.

Sampling all binding modes within a single free energy calculation is usually impractical without some form of enhanced sampling or at least Hamiltonian replica exchange~\cite{Wang:2013:JournalofComputer-AidedMolecularDesign} because barriers for binding mode interconversion result in kinetics which are too slow compared to simulation timescales~\cite{Mobley:2006:TheJournalofChemicalPhysics, Palma:2012:J.Comput.Chem., Mobley:2012:TheJournalofChemicalPhysics, Gill:2018:J.Phys.Chem.B}.
Hamiltonian exchange, coupled with appropriate restraints, can allow the ligand to relatively rapidly exchange between potential binding modes when non-interacting, accelerating sampling of binding modes~\cite{Wang:2013:JournalofComputer-AidedMolecularDesign}.
\todo[inline, color={yellow!40}]{DLM probably need more background refs on Hamiltonian lambda exchange here.}

Separation of states provides a simple though potentially expensive alternative, where each stable binding mode is considered separately with a binding free energy calculation restricted to that binding mode, and then (as long as the binding modes are non-overlapping) the resulting component binding free energies can be combined into a total~\cite{Mobley:2006:TheJournalofChemicalPhysics, Mobley:2012:TheJournalofChemicalPhysics}.
This approach necessitates a separate binding free energy calculation for each potential binding mode, however, so it can be computationally quite costly.
If relative populations of different stable binding modes were available from some other technique, it could make this separation of states approach considerably more efficient~\cite{Mobley:2012:TheJournalofChemicalPhysics, Gill:2018:J.Phys.Chem.B}.

@@

To deal with this, some practitioners have actually computed relative binding free energies of different binding modes of the same ligand~\cite{Palma:2012:J.Comput.Chem.} [ref Jorgensen (?) etc.].
For example, a mutation which adds a methyl to an aromatic ring of a larger ligand might yield one result if the methyl points in one direction, and a different value if it points in the other due to slow ring motions. [e.g. get ref from Sukanya]
One could compute the free energy of turning off the methyl group in one orientation and turning it back on in the other orientation to obtain the free energy difference between the two potential binding modes.
While this approach has precedent, it is relatively difficult to automate at present and requires considerable care.

@@

%%%%%%%%%%%%%%%%%%%%
%              Terminology                  %
%%%%%%%%%%%%%%%%%%%%
\section{Terminology and abbreviations}
\label{sec:tem-abbrev}
\begin{itemize}
\item Feature Box covering major technical terms and abbreviations
\item Examples:
\begin{itemize}
\item EXP, BAR, MBAR
\item Double decoupling, single-topology, dual-topology, hybrid-topology, coupled-topology
\item FEP (free energy peturbation), alchemical, AFE (alchemical free energy)
\end{itemize}
\end{itemize}
%%%%%%%%%%%%%%%%%%%%
%                Software                    %
%%%%%%%%%%%%%%%%%%%%
\section{Available software -- a summary}
\todo[inline]{Samar: write section}
\label{sec:software}
There are several softwares availble for carrying out alchemical free energy calculations. In this section we discuss some of the popular commercial and free tools along with their features. 
\begin{itemize}
\item Commercial:
   \begin{itemize}
    \item FEP+ is a tool offered by Schrodinger Inc. under a commercial license. It has an intuitive GUI which makes it easier for non-experts to run alchemical free energy calculations and analyze the results. It runs DESMOND MD package under the hood and hence parallelizes very well on the GPUs. 
    \end{itemize}
\item Free or low-cost for academics / commercial for industry:
	\begin{itemize}
	\item CHARMM has a variety of tools developed over the years. PERT module can be used to define initial and final states and define the intermediate lambda points. FREN and BAR modules can be used to analyze the data after the MD run. Lambda-dynamics based free energy calulation can be carried out using the BLOCK module.  
	\item TIES and AMBER FEW? (Peter Coveney)
	\item AMBER, including its new pmemd.cuda version support free energy calculation. 
	\end{itemize}
\item Free (libre) open source:
	\begin{itemize}
	\item PLUMED is an open source tool which enables the usage of a variety of MD engines. It is designed as a plugin for MD packages such that it analyzes the trajectory on the fly. It also offers a VMD based plugin for the computation of collective variables.   	
	\item SIRE  is a free, open source, multiscale molecular simulation framework, written to allow computational modellers to quickly prototype and develop new algorithms for molecular simulation and molecular design. 
	\item YANK is a tool developed by John Chodera and group on the top of OpenMM MD package. It allows the users to write their inputs in easy-to-use YAML format.
	\item GROMACS is a molecular simulation package with a significant number of free energy methods implementations. The LiveCOMS GROMACS tutorial has an example free energy calculation~\cite{Lemkul2018From}.
	\item pmx for mutations
	\end{itemize}
\item Setup tools
	\begin{itemize}
	\item FESetup: AMBER, gromacs, Sire
	\item Lomap/Lomap2 : Relative alchemical transformation graph planning
	\item CHARMM-GUI is a web based tool for setting up a variety of MD simulations. It can be used to generate CHARMM sctipts for solvation and ligand-binding free energy calculations.
	\end{itemize}
\item Analysis tools:
	\begin{itemize}
	\item Free Energy Workflows: Sire-specific free energy map analysis using weighted path averages
	\url{https://github.com/michellab/freenrgworkflows}
	\item Alchemlyb: Multipackage free energy analysis
	\url{https://github.com/alchemistry/alchemlyb}
	\item pymbar: MBAR implementation, but have to roll your own analysis wrapper
	\url{https://github.com/choderalab/pymbar}
	\end{itemize}
\end{itemize}

>>>>

Below, we provide a non-exhaustive list of topics we have not addressed with some references to provide starting points on these more advanced topics:
\begin{itemize}
\item Covalent inhibition.
\item Complex association stoichiometry (beyond simple 1:1 association)~\cite{awesome reference}. 
\item Endpoint free energy methods such as MM-PBSA~\cite{genheden2015mm} and LIE~\cite{gutierrez-de-teran2012linear}
\item Free energy methods that extract the ligand using geometric order parameters and potential of mean force methods~\cite{heinzelmann2017attachpullrelease}
\item Forcefield dependence for protein, ligand, ions, co-solvents, and co-factors. Different studies have looked at the influence of force fields and it is assumed the user has made an adequate choice for the system under study.~\cite{loeffler2018reproducibility, vassetti2019assessment, lopes2015current} 
\end{itemize}

@@

How can we compute $\Delta G_{\mathrm{bind}}$ from a molecular dynamics simulation? Running a direct simulation of the binding events is too computationally costly, meaning we cannot easily estimate the configurational integrals of the bound and unbound state from the simulation. This is where alchemical free simulations come in. We can write down a thermodynamic cycle as seen in ~\ref{fig:fig3_binding-thermodynamic-cycle}, for a relative free energy simulation. Based on this cycle we can compute the relative free energy of binding between two compounds $A$ and $B$, by going along the arrow of $\Delta G_{\mathrm{bound}}$, where we introduce a second compound that is similar to the first one. Now $\Delta G_{\mathrm{bound}}$ can be expressed as a ratio of partition functions of toluene (A) and benzyl alcohol (B). 

@@

Current alchemical free energy calculations seem to achieve, in favorable cases, RMS errors around 1-2 kcal/mol depending on force field, system, and a variety of other factors such as simulation time, sampling method, and whether the calculations employed are absolute or relative. An overview of benchmark datasets and case studies can be found in section \ref{sec:benchmark} at the end of this document.
\todo[inline,color={green!20}]{JDC: We should probably provide a short list of key benchmark references here, broken down by force field, and keep this up to date. GT: would a table, at the end of the document be an easier way to do this, only requiring some brief comments in the text? ASJSM: I like the table idea. MK: Feel free to remove this todo item if you are happy with the table provided. I refrained from just listing lots of case studies, as I think we should highlight the recommended datasets which should be used for benchmarking. Nevertheless, I have cited an review article containing such a summary as well as a recent publication comparing FEP+ and Gromacs in the brief text below the table.}
However, the domain of applicability is a significant concern~\cite{sherborne2016collaborating, cournia2017relative}, especially for relative calculations, which typically require a bound structure of a closely
related ligand as a starting point. Additional factors such as slow protein or ligand rearrangements, uncertainties in ligand binding mode, or charged ligands can make these calculations far less reliable and more of a research effort.

@@

If the binding site location is uncertain but the structure of the receptor is well defined and plausible binding sites are identified, you are more likely to chose an absolute free energy protocol to compute the standard free energy of binding of the ligand to a set of binding sites. This requires the user to prepare input files describing the bound conformation in different putative binding sites (REF). The apparent binding free energy of the ligand may be obtained by combining the individual binding site free energies, which also indicate where the ligand is more likely to bind. In this case using a docking program to generate initial structures is the way to go. Different commercial and none commercial tools are available, such as: rDock, Autodock Vina, Glide, or Flare to name a few~\cite{ruiz-carmona2014rdock, trott2010autodock, friesner2004glide, cheeseright2006molecular}. 
If the putative binding sites are not apparent because of for instance significant induced-fit effects it may be challenging to obtain meaningful free energies of binding. One would have to account for the free energy change for forming a binding site in the target receptor. However, it may still be possible to obtain useful binding free energy estimates that may be compared between different ligands.

@@

\paragraph{Topologies} A critical first step in relative calculations is to select an approach to these calculations, determining whether to use a \emph{dual topology}, \emph{single topology}, or \emph{hybrid topology} approach to relative calculations.
\begin{figure}
    \includegraphics[width=0.95\columnwidth]{paper/figures/fig4_topol/Fig4.pdf}
    \caption{top: single topology bottom: dual topology Figure adapted from \url{http://www.alchemistry.org/wiki/Constructing_a_Pathway_of_Intermediate_States}}
    \label{fig:fig4_topology}
\end{figure} 
The distinction between these can be illustrated by considering a hypothetical transformation from molecule A to molecule B, where both atoms share a common substructure but differ in which functional groups are present; e.g. consider a transformation of benzene (A) to benzyl alcohol (B)~\ref{fig:fig4_topology}.
In this case the common substructure is the benzene ring, though perhaps may be larger depending on how it is defined, as we discuss below.
In single topology calculations, the overall transformation is set up to involve as few additional atoms as possible, so benzene would be typically changed into benzyl alcohol by changing one of the hydrogens into a carbon. To this new carbon you also find non-interacting atoms called ``dummy atoms'' (retaining their bonded interactions but not interacting with the rest of the system), which will become the two hydrogens attached to the new carbon and the OH group. Bond parameters as well as partial charges are adjusted accordingly. 
Thus in a single topology calculation, atoms may change their type so relatively few dummy atoms are created. This is illustrated in fig.~\ref{fig:fig4_topology} (a). 
In contrast, in a dual topology free energy calculation, no atoms are allowed to change type~\cite{shirts2012best}. This means that the benzene to benzyl alcohol transformation involves starting with benzene plus the non-interacting dummy atoms making up the hydroxy methyl group, then passing through an intermediate state where atoms which are becoming dummy atoms or ceasing being dummy atoms are partially interacting (this state may or may not be well defined~\cite{mobley2014blind}  and culminating in a state where benzyl alcohol is present along with the additional dummy atoms which was previously a corresponding hydrogen of the benzene. Fig.~\ref{fig:fig4_topology} (b) depicts how such a dual topology works.

@@

There are different tools that allow the generation of MCSS matches as well as single topology input. A large number of software tools are nowadays available for computing MCSS matches, some of the rely on RDKit for this computation, such as LOMAP or FESetup and partially BioSimSpace, others such as fkckombu are standalone tools~\cite{rdkit2019Dec,loeffler2015fesetup, liu2013lead, hedges2019biosimspace, kawabata20143d}. Schr\"{o}dinger's FEP+ planning tool is based on a version of LOMAP, also allows MCSS matching and planning single topology calculations between molecules~\cite{wang2015accurate}.

@@

In binding free energy calculations, extra care should be taken when simulating the bound state, especially when dealing with weak binders and absolute free energy calculations in combination with enhanced sampling techniques.
In principle, only configurations of the receptor-ligand complex that we consider "bound" should be sampled from the bound state, and, in atomistic simulations, this requires to establish a structural definition of the bound state.
In practice, the simulation of the bound state starts with the ligand already placed in the binding site and relies on kinetic trapping to maintain a bound complex.
However, this strategy may not be sufficient if the complex dissociation rate is high or if methodologies such as Hamiltonian replica exchange~\cite{bitetti-putzer2003generalized} and expanded ensemble~\cite{chodera2011replica} are employed in absolute free energy calculations to enhance sampling since, in both cases, the ligand may find a way out of the binding site on timescales that are achievable by modern MD simulations.
A solution commonly adopted in these cases is the use of one or more restraints making the unbound configurations energetically unfavorable through the addition of extra terms in the potential function.
In practice, even with restraints, it is not always trivial to force a molecular dynamics simulation to explore a restricted region of the configurational space with a complicated geometry such as a binding site.
If the ligand is a tight binder\todo[inline, color={red!40}]{AR: Should we give an order of magnitude to define a tight and a weak binder?}, it is usually safe to employ a restraint that allows some of the unbound configurations to be sampled since they generally contribute negligibly to the partition function (i.e., they have a relatively small Boltzmann weight) as long as the sampled volume is not so large that their cumulative contribution becomes significant.
However, this is more problematic when dealing with weak binders as the unbound configurations can have a non-negligible Boltzmann weight, and their binding affinity can exhibit a significant dependency on the restraint type and parameters that are used to determine the sampled volume~\cite{}.
Moreover, it should be noted that the additional potential energy terms used to model the restraints can introduce a bias in the predicted free energy.
The bias can be removed at the analysis stage through reweighting techniques, but this procedure can increase the statistical uncertainty of the binding free energy estimate when the restraints are so strong that the overlap with the reweighted state is diminished~\cite{}.

@@

In a discovery setting, new ligands typically have unknown or at least uncertain binding modes~\cite{kaus2015how, plountprice2000analysis,mobley2009binding,calabro2016elucidation},  complicating binding free energy estimation.
This uncertainty is because it is usually not desirable to estimate a binding affinity for a ligand which already has an available bound structure, since such a compound has already been tested.
To deal with prospective ligands with unknown binding modes, discovery projects commonly assume that modifications of functional groups on a common scaffold result in a consistent binding mode across all members of a series.
This is not necessarily always the case~\cite{kaus2015how}, as reviewed elsewhere~cite{mobley2009binding} and in some cases unexpected binding mode changes can be the origin of apparent non-additivity in structure-activity relationships~\cite{calabro2016elucidation}.
Binding modes also tend to be particularly variable in the case of fragments, which often may have multiple relevant binding modes~\cite{}.

@@

In some cases, it is tempting to simply use docking techniques to generate initial bound structures for starting molecular dynamics simulations.
However, timescales for binding mode interconversion are usually slow compared to MD/free energy timescales, meaning that simulations started from different potential binding modes are likely to yield disparate computed binding free energies~\cite{mobley2006use, palma2012computation, mobley2012perspective, gill2018binding} .
And docking techniques are good at identifying sterically reasonable potential binding modes, but still perform relatively poorly at identifying a single dominant binding mode \emph{a priori}.~\cite{}

It is worth highlighting a recent SAMPL blind challenge on HIV integrase as an illustration of this. 
Many submissions, using state-of-the-art methods, had difficulty even predicting which \emph{binding site} ligands would bind in (most submissions placed more than half of the ligands into the incorrect binding site), and even given correct binding sites, the binding mode within each site was also quite difficult to predict~\cite{mobley2014blind}.
The best performing submission for predicting binding modes actually ended up being a human expert (aided by computational tools) with more than 10 years of experience on the particular target~\cite{voet2014combining}, rather than a fully automated approach.
While free energy calculations on this set had some success, many of the failures actually ended up being cases where the binding mode selected as input for free energy calculations was later found to be incorrect~\cite{gallicchio2014virtual}, highlighting the importance of these issues.

One approach which has shown some success is to retain diverse potential binding modes from docking, perform short MD simulations of these to identify distinct stable binding modes, and then consider these in subsequent calculations~\cite{gallicchio2014virtual, mobley2006use,rocklin2013blind, boyce2009predicting, mobley2007predicting}.

Routes to handle multiple potential binding modes are different depending on whether absolute or relative calculations are selected, unless a method is available to estimate the relative populations of different stable binding modes in advance (e.g. such as the BLUES approach currently in development~\cite{gill2018binding}), in which case this approach could be applied to assist both types of calculations.

\paragraph{Handling multiple potential binding modes within absolute calculations.}
Within absolute binding free energy calculations, multiple potential binding modes can be handled by two main strategies: Consider each binding mode separately (a separation of states strategy) or sample all binding modes within a single simulation~\cite{mobley2012perspective}.
This couples to the choice of restraints selected, as some restraints will allow transitions between binding modes and even binding sites (Section~\ref{sec:standardstate-restraints}), and others do not.

Sampling all binding modes within a single free energy calculation is usually impractical without some form of enhanced sampling or at least Hamiltonian replica exchange~\cite{wang2013identifying} because barriers for binding mode interconversion result in kinetics which are too slow compared to simulation timescales~\cite{mobley2006use, palma2012computation,mobley2012perspective, gill2018binding}.
Hamiltonian exchange, coupled with appropriate restraints, can allow the ligand to relatively rapidly exchange between potential binding modes when non-interacting, accelerating sampling of binding modes~\cite{wang2013identifying}.
\todo[inline, color={yellow!40}]{DLM probably need more background refs on Hamiltonian lambda exchange here.}

Separation of states provides a simple though potentially expensive alternative, where each stable binding mode is considered separately with a binding free energy calculation restricted to that binding mode, and then (as long as the binding modes are non-overlapping) the resulting component binding free energies can be combined into a total~\cite{mobley2006use, mobley2012perspective}.
This approach necessitates a separate binding free energy calculation for each potential binding mode, however, so it can be computationally quite costly.
If relative populations of different stable binding modes were available from some other technique, it could make this separation of states approach considerably more efficient~\cite{mobley2012perspective, gill2018binding}.

@@

To deal with this, some practitioners have actually computed relative binding free energies of different binding modes of the same ligand~\cite{palma2012computation} [ref Jorgensen (?) etc.].
For example, a mutation which adds a methyl to an aromatic ring of a larger ligand might yield one result if the methyl points in one direction, and a different value if it points in the other due to slow ring motions. [e.g. get ref from Sukanya]
One could compute the free energy of turning off the methyl group in one orientation and turning it back on in the other orientation to obtain the free energy difference between the two potential binding modes.
While this approach has precedent, it is relatively difficult to automate at present and requires considerable care.

@@

%%%%%%%%%%%%%%%%%%%%
%              Terminology                  %
%%%%%%%%%%%%%%%%%%%%
\section{Terminology and abbreviations}
\label{sec:tem-abbrev}
\begin{itemize}
\item Feature Box covering major technical terms and abbreviations
\item Examples:
\begin{itemize}
\item EXP, BAR, MBAR
\item Double decoupling, single-topology, dual-topology, hybrid-topology, coupled-topology
\item FEP (free energy peturbation), alchemical, AFE (alchemical free energy)
\end{itemize}
\end{itemize}
%%%%%%%%%%%%%%%%%%%%
%                Software                    %
%%%%%%%%%%%%%%%%%%%%
\section{Available software -- a summary}
\todo[inline]{Samar: write section}
\label{sec:software}
There are several softwares availble for carrying out alchemical free energy calculations. In this section we discuss some of the popular commercial and free tools along with their features. 
\begin{itemize}
\item Commercial:
   \begin{itemize}
    \item FEP+ is a tool offered by Schrodinger Inc. under a commercial license. It has an intuitive GUI which makes it easier for non-experts to run alchemical free energy calculations and analyze the results. It runs DESMOND MD package under the hood and hence parallelizes very well on the GPUs. 
    \end{itemize}
\item Free or low-cost for academics / commercial for industry:
	\begin{itemize}
	\item CHARMM has a variety of tools developed over the years. PERT module can be used to define initial and final states and define the intermediate lambda points. FREN and BAR modules can be used to analyze the data after the MD run. Lambda-dynamics based free energy calulation can be carried out using the BLOCK module.  
	\item TIES and AMBER FEW? (Peter Coveney)
	\item AMBER, including its new pmemd.cuda version support free energy calculation. 
	\end{itemize}
\item Free (libre) open source:
	\begin{itemize}
	\item PLUMED is an open source tool which enables the usage of a variety of MD engines. It is designed as a plugin for MD packages such that it analyzes the trajectory on the fly. It also offers a VMD based plugin for the computation of collective variables.   	
	\item SIRE  is a free, open source, multiscale molecular simulation framework, written to allow computational modellers to quickly prototype and develop new algorithms for molecular simulation and molecular design. 
	\item YANK is a tool developed by John Chodera and group on the top of OpenMM MD package. It allows the users to write their inputs in easy-to-use YAML format.
	\item GROMACS is a molecular simulation package with a significant number of free energy methods implementations. The LiveCOMS GROMACS tutorial has an example free energy calculation~\cite{lemkul2018From}.
	\item pmx for mutations
	\end{itemize}
\item Setup tools
	\begin{itemize}
	\item FESetup: AMBER, gromacs, Sire
	\item Lomap/Lomap2 : Relative alchemical transformation graph planning
	\item CHARMM-GUI is a web based tool for setting up a variety of MD simulations. It can be used to generate CHARMM scripts for solvation and ligand-binding free energy calculations.
	\end{itemize}
\item Analysis tools:
	\begin{itemize}
	\item Free Energy Workflows: Sire-specific free energy map analysis using weighted path averages
	\url{https://github.com/michellab/freenrgworkflows}
	\item Alchemlyb: Multipackage free energy analysis
	\url{https://github.com/alchemistry/alchemlyb}
	\item pymbar: MBAR implementation, but have to roll your own analysis wrapper
	\url{https://github.com/choderalab/pymbar}
	\end{itemize}
\end{itemize}
